Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in early 2020: An exploratory analysis

### 2021-06-04 manuscript/hmm.Rmd

>>>>

```{r}
# Estimates for FPV for metaanalysis
hmm_fit <- c_trt_m3_fw_fit
numerator_serie_data <- serie_data_28_fw %>% mutate(took_favipiravir = 1)
denominator_serie_data <- serie_data_28_fw %>% mutate(took_favipiravir = 0)

  combined_data <- hmm_fit$data
  combined_data$serie_data = rbind(
    numerator_serie_data %>% mutate(.serie = paste0("__num_", .serie), is_numerator = TRUE),
    denominator_serie_data %>% mutate(.serie = paste0("__denom_", .serie), is_numerator = FALSE)
  ) %>%
    mutate(.serie = factor(.serie))

  combined_data$initial_states = rep(hmm_fit$data$initial_states, 2)

  oxygen_states_ids <- which(hmm_fit$data$hidden_state_data$corresponding_obs == "Oxygen")
  ventilation_states_ids <- which(hmm_fit$data$hidden_state_data$corresponding_obs == "Ventilated")

  numerator_indices <- 1:length(hmm_fit$data$initial_states) + length(hmm_fit$data$initial_states)
  denominator_indices <- 1:length(hmm_fit$data$initial_states)

  numerator_indices_oxygen <- numerator_indices[hmm_fit$data$initial_states == "AA_worsening"]
  denominator_indices_oxygen <- denominator_indices[hmm_fit$data$initial_states == "AA_worsening"]

  numerator_indices_ventilated <- numerator_indices[hmm_fit$data$initial_states != "Ventilated_worsening"]
  denominator_indices_ventilated <- denominator_indices[hmm_fit$data$initial_states != "Ventilated_worsening"]

states_rep <- 4
states_list <- list()
for(i in 1:states_rep) {
  states_list[[i]] <- posterior_epred_rect(hmm_fit, newdata = combined_data, method = posterior_epred_rect_simulate, cores = 6)
}

states <- do.call(abind::abind, args = c(states_list, list(along = 3)))

  needed_ventilation <- apply(states, MARGIN = c(2,3), FUN = function(x) { any(x %in% ventilation_states_ids) })
  needed_oxygen <- apply(states, MARGIN = c(2,3), FUN = function(x) { any(x %in% oxygen_states_ids) })

  samples_oxygen_probs_numerator <- 
    colMeans(needed_oxygen[numerator_indices_oxygen, ])
  samples_oxygen_probs_denominator <- 
    colMeans(needed_oxygen[denominator_indices_oxygen, ])

  log_oxygen_odds_numerator <- log(samples_oxygen_probs_numerator) - log1p(-samples_oxygen_probs_numerator)
  log_oxygen_odds_denominator <- log(samples_oxygen_probs_denominator) - log1p(-samples_oxygen_probs_denominator)

  log_OR_oxygen <- log_oxygen_odds_numerator - log_oxygen_odds_denominator

    samples_ventilation_probs_numerator <- 
    colMeans(needed_ventilation[numerator_indices_ventilated, ])
  samples_ventilation_probs_denominator <- 
    colMeans(needed_ventilation[denominator_indices_ventilated, ])

  log_ventilation_odds_numerator <- log(samples_ventilation_probs_numerator) - log1p(-samples_ventilation_probs_numerator)
  log_ventilation_odds_denominator <- log(samples_ventilation_probs_denominator) - log1p(-samples_ventilation_probs_denominator)

  log_OR_ventilation <- log_ventilation_odds_numerator - log_ventilation_odds_denominator

  # There is a very tiny fraction of NaNs where no patient is predicted to need ventilation. remove those.
  log_OR_ventilation <- log_OR_ventilation[is.finite(log_OR_ventilation)]

  death_res <- hmm_hypothesis_res_list[["c_trt_rate_m3_fw"]] %>% filter(hypothesis == "favipiravir_death")

res_fpv <- tibble::tribble(
  ~event, ~estimate, ~se, ~`lower95`, ~`upper95`,
  "Death",  death_res$point_estimate, NA, death_res$ci_low, death_res$ci_high,
  "Oxygen", mean(log_OR_oxygen), sd(log_OR_oxygen), quantile(log_OR_oxygen, 0.025), quantile(log_OR_oxygen, 0.975),
  "Ventilated", mean(log_OR_ventilation), sd(log_OR_ventilation), quantile(log_OR_ventilation, 0.025, na.rm=TRUE), quantile(log_OR_ventilation, 0.975, na.rm=TRUE)
  )

write_csv(res_fpv, file = here::here("local_temp_data", "fpv_meta","main_events.csv"))
```

### 2021-06-04 manuscript/manuscript.Rmd

---
title: "Detailed disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic: An exploratory analysis" 
output: 
  #rticles::arxiv_article
  #pdf_document
  # bookdown::pdf_document2:
  #    toc: false
  bookdown::word_document2: 
    toc: false
    number_sections: false
    reference_docx: "output_template.docx"
abstract: |  
  **Abstract:** We collected a multi-centric retrospective dataset of patients (N = 213) who were admitted to ten hospitals in Czech Republic and tested positive for SARS-CoV-2. The dataset contains baseline patient characteristics, breathing support required, pharmacological treatment received and multiple markers on daily resolution. Patients in the dataset were treated with hydroxychloroquine (N = 108), azithromycin (N = 72), favipiravir (N = 9), convalescent plasma (N = 7), dexamethasone (N = 4) and remdesivir (N = 3), often in combination. Most patients were admitted during the first wave of the epidemic. To explore association between treatments and patient outcomes we performed multiverse analysis, observing how the conclusions change between defensible choices of statistical model, predictors included in the model and other analytical degrees of freedom.  Weak evidence to constrain the potential efficacy of azithromycin and favipiravir can be extracted from the data. Additionally, we performed external validation of several proposed prognostic models for Covid-19 severity showing that they mostly perform unsatisfactorily on our dataset.
bibliography: Covid19.json
csl: journal-of-medical-virology.csl
#date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
author: "`r { source(here::here('manuscript', 'authors.R')); get_authors_string() }`"
---

@@

A convenience sample of patients from `r n_hospitals` sites was collected. For each site, the dataset contains all patients hospitalized in the participating wards over the data collection period. The data collection started at the onset of the Covid-19 pandemic (except for one site where some older records were inaccessible), but the end date for collection differs between sites due to time constraints of the participating physicians. Three sites included total of `r n_second_wave` patients that could be considered part of "second wave", i.e. admitted after September 1st, last patient included in the dataset was admitted on October 12th. See Supplementary Figure 1 for per-site data collection periods. Patients over the age of 18 were included if they had PCR-confirmed infection of SARS-CoV-2 and were not participating in a clinical trial of any Covid-19 pharmacotherapy.

>>>>

---
title: "Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in early 2020: An exploratory analysis" 
output: 
  #rticles::arxiv_article
  #pdf_document
  # bookdown::pdf_document2:
  #    toc: false
  bookdown::word_document2: 
    toc: false
    number_sections: false
    reference_docx: "output_template.docx"
abstract: |  
  **Abstract:** We collected a multi-centric retrospective dataset of patients (N = 213) who were admitted to ten hospitals in Czech Republic and tested positive for SARS-CoV-2 during the early phases of the pandemic in March - October 2020. The dataset contains baseline patient characteristics, breathing support required, pharmacological treatment received and multiple markers on daily resolution. Patients in the dataset were treated with hydroxychloroquine (N = 108), azithromycin (N = 72), favipiravir (N = 9), convalescent plasma (N = 7), dexamethasone (N = 4) and remdesivir (N = 3), often in combination. Most patients were admitted during the first wave of the epidemic. To explore association between treatments and patient outcomes we performed multiverse analysis, observing how the conclusions change between defensible choices of statistical model, predictors included in the model and other analytical degrees of freedom.  Weak evidence to constrain the potential efficacy of azithromycin and favipiravir can be extracted from the data. Additionally, we performed external validation of several proposed prognostic models for Covid-19 severity showing that they mostly perform unsatisfactorily on our dataset.
bibliography: Covid19.json
csl: journal-of-medical-virology.csl
#date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
author: "`r { source(here::here('manuscript', 'authors.R')); get_authors_string() }`"
---

@@

A convenience sample of patients from `r n_hospitals` sites was collected. For each site, the dataset contains all patients hospitalized in the participating wards over the data collection period. The data collection started at the onset of the Covid-19 pandemic in March 2020 (except for one site where some older records were inaccessible), but the end date for collection differs between sites due to time constraints of the participating physicians. Three sites included total of `r n_second_wave` patients that could be considered part of "second wave", i.e. admitted after September 1st, last patient included in the dataset was admitted on October 12th. See Supplementary Figure 1 for per-site data collection periods. Patients over the age of 18 were included if they had PCR-confirmed infection of SARS-CoV-2 and were not participating in a clinical trial of any Covid-19 pharmacotherapy.
